### 2016-06-01 paper/presentation.tex
presentation

\documentclass[11pt]{beamer}

%% Sprache und Zeichensatz:
\usepackage[applemac]{inputenc}
%\usepackage[ngerman]{babel}
\usepackage[T1]{fontenc}
\usepackage[ngerman,english]{babel}
\usepackage{lmodern}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amstext}
%\usepackage{amsthm}

\usepackage{setspace}

\usetheme{boxes}
\usepackage{dsfont}

\usecolortheme{rose}
\useinnertheme[shadow]{rounded}

%---------------------------------------------------------------------------------------------
%
%%Pakete Mathematik
%\usepackage{amsmath, amssymb, amstext, amsfonts, mathrsfs}
%\usepackage[squaren]{SIunits}

%---------------------------------------------------------------------------------------------
%% Für eine Titelseite:

%%%%%%

\usepackage[matrix,arrow]{xy}
\usepackage{stmaryrd}

\DeclareMathOperator*{\esssup}{ess.sup}
\DeclareMathOperator*{\defgl}{:=}
\DeclareMathOperator*{\defgr}{=:}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\3}{ ^{3\times3} }
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\M}{\mathbb{M}}
\newcommand{\Sym}{\mathbb{S}^3}
\newcommand{\Mpos}{\M_+^3}
\newcommand{\Orth}{\mathbb{O}^3}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left |#1\right |}
\newcommand{\ub}{\bar{u}}
\newcommand{\yb}{\bar{y}}
\newcommand{\scprod}[2]{\langle#1 , #2 \rangle}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\G}{\mathcal{G}}

% TIKZ DIAGRAMS

\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}

\tikzset{
    block/.style={
           rectangle,
           rounded corners,
           draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
     block2/.style={
           %rectangle,
           %rounded corners,
           fill=black!20,
           %draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
     frameless/.style={
           %rectangle,
           %rounded corners,
           %draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
}

\usepackage{tikz-cd}
\tikzset{
  shift left/.style ={commutative diagrams/shift left={#1}},
  shift right/.style={commutative diagrams/shift right={#1}}
}

%%%%%%

\title{High Performance Computing Project\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration} 
\author{Manuel Demmeler}
\institute{Technische Universität München}
\date{}

\AtBeginSection[]{ 
	\frame {
		%\frametitle{Gliederung}
		\tableofcontents[current, currentsubsection]
	} 
}

\begin{document}

\frame{\titlepage}

%\begin{frame}
%	\frametitle{Gliederung}
%	\tableofcontents
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%
%\section{Motivation}
\begin{frame}
\frametitle{Motivation}
	\begin{center}
		\includegraphics[height=4cm]{configspacetrafo.png}
	\end{center}
	\begin{itemize}
		\item Problem: find a feasible trajectory of a robot through a setting of obstacles
		\item Often necessary to pass feasible initial tajectories to optimization algorithms
	\end{itemize}
\end{frame}

\newtheorem*{defi}{Definition}

\begin{frame}
\frametitle{Problem Statement}
	\begin{itemize}
		\item Obstacles can not be assumed to have a closed form
		\item Therefore they are defined by an indicator function
		\begin{equation*}
			I: \Omega \subset \R^d \rightarrow \{ 0,1 \}
		\end{equation*}
		\item Collision for $q \in \Omega$, $I(q) = 1$ 
	\end{itemize}

\end{frame}

\begin{frame}
\frametitle{PRM Algorithm}
	PRM: Grow graphs from given start and endpoints $q_s, q_e \in \Omega$ until a connection is found, by repeating
	\begin{itemize}
		\item sample new node $v$ randomly in environment of old nodes
		\item check for all possible neighbors $w$, if the linear connection is feasible, in this case add edge $\{v,w\}$ to the graph
		\item break, if the two graphs have been connected.
	\end{itemize}
	\begin{defi}
	For two points \(v\) and \(w\) we define the connection with stepsize \(h>0\) as
		\begin{equation*}
			[v,w]_h \defgl \{ q = \lambda v+(1-\lambda) w\ |\ \lambda \in [0,1], \norm{q-v} \in \N_0 h \}.
		\end{equation*}
		We say, \(q_1\) and \(q_2\) are connected with stepsize \(h>0\), if
		\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
	\end{defi}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%
%\section{Robotics Application}

\begin{frame}
\frametitle{Robotics Application}
	\begin{center}
		\includegraphics[height=3cm]{configspacetrafo.png}
	\end{center}
	\begin{itemize}
		\item For each linear connection, every intermediate configuration has to be checked for collisions
		\item $\Rightarrow$ Many independent and similar kinematics and collision calculations
		\item $\Rightarrow$ Well suited for usage of GPU
		\item Every GPU thread can make one intermediate test
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Indicator Function Implementation}
\begin{itemize}
	\item Input: vectors of start and end nodes for each potential edge
	\item Output: vector of feasible edges
	\item CPU determines number of threads needed for each edge
		 and passes nodes to GPU 
	\item Each CUDA thread determines his intermediate configuration and checks for collsion
	\begin{itemize}
		\item Kinematics with Denavit-Hartenberg transformations
		\item Geometry parts are convex polytopes
		\item Collision detection: Chung-Wang Collision Algorithm
	\end{itemize}
\end{itemize}
\end{frame}

%\section{Parallelization}

\begin{frame}
\frametitle{Project Overview}
\begin{center} 
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[auto]%node distance=0.5cm and 1.8cm]
	\node[block] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver}\\
		generate nodes\\
		store graph
		\end{tabular}
	};
	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace}\\
		indicator fct. \\
		interface
		\end{tabular}
	};
	\node [block, right=of config] (robo) {
		\begin{tabular}{c}
		\textbf{RobotConfigspace}\\
		robot kinematics\\
		collision detection
		\end{tabular}
	};
	\node [block, right=of robo] (gpu) {\textbf{GPU}};

	\path[-stealth,thick, shift left=1.0ex]	(prm) edge (config) ;
	\path[-stealth,thick, shift left=1.0ex]	(config) edge  (prm) ;
	\path[-stealth,thick, shift left=1.0ex]	(config) edge (robo);
	\path[-stealth,thick, shift left=1.0ex]	(robo) edge  (config);
	\path[-,thick]					(robo) edge (gpu);
\end{tikzpicture}
}
\end{center}
	\begin{itemize}
	\item PRM algorithm and Configuration space independent from each other
	\item Only connected through indicator function represented by Configspace interface
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Second Level of Parallelization}
	Version 1:
	\begin{itemize}
	\item Use multiple GPUs parallely
	\item Sample and connect new nodes on each processor
	\item Exchange nodes and edges
	\end{itemize}

\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
	\node[block, minimum width=3cm] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver 1}
		\end{tabular}
	};

	\node[frameless, below=of prm, minimum width=3cm] (space) {$\vdots$};

	\node[block, below=of space, minimum width=3cm] (prm2) {
		\begin{tabular}{c}
		\textbf{PRMSolver n}
		\end{tabular}
	};

	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace 1}
		\end{tabular}
	};
	\node[frameless, below=of config] (space2) {$\vdots$};
	\node [block, below=of space2] (config2) {
		\begin{tabular}{c}
		\textbf{Configpace n}
		\end{tabular}
	};

	\node [block, right=of config] (gpu) {\textbf{GPU 1}};
	\node [block, right=of config2] (gpu2) {\textbf{GPU n}};

	\path[-latex',thick, shift left=1.0ex]	(prm) edge (config) 
								(config) edge (prm);
	\path[-,thick]					(config) edge (gpu);

	\path[-latex',thick, shift left=1.0ex]	(prm2) edge (config2) 
								(config2) edge (prm2) ;
	\path[-,thick]					(config2) edge (gpu2);

	\path[-latex',thick, shift left=1.0ex]	(prm) edge (space)
								(space) edge (prm)
								(prm2) edge (space)
								(space) edge (prm2) ;

\end{tikzpicture}
}
\end{center}

\end{frame}

\begin{frame}
\frametitle{Second Level of Parallelization}
	Version 2:
	\begin{itemize}
		\item Grow whole graphs on every processor
		\item Exchange only edges
	\end{itemize}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
	\node[block, minimum width=3cm] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver 1}
		\end{tabular}
	};

	\node[frameless, below=of prm, minimum width=3cm] (space) {$\vdots$};

	\node[block, below=of space, minimum width=3cm] (prm2) {
		\begin{tabular}{c}
		\textbf{PRMSolver n}
		\end{tabular}
	};

	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace 1}
		\end{tabular}
	};
	\node[frameless, below=of config] (space2) {$\vdots$};
	\node [block, below=of space2] (config2) {
		\begin{tabular}{c}
		\textbf{Configpace n}
		\end{tabular}
	};

	\node [block, right=of config] (gpu) {\textbf{GPU 1}};
	\node [block, right=of config2] (gpu2) {\textbf{GPU n}};

	\path[-latex',thick, shift left=1.5ex]	(prm) edge (config) ;
	\path[-,thick]					(config) edge (gpu);

	\path[-latex',thick, shift right=1.5ex]	(prm2) edge (config2) ;
	\path[-,thick]					(config2) edge (gpu2);

	\path[-latex',thick]				(config.west) edge  (prm.east) 
								(config.west) edge (space.east)
								(config.west) edge (prm2.east)
								(config2.west) edge  (prm.east) 
								(config2.west) edge (space.east)
								(config2.west) edge (prm2.east);

\end{tikzpicture}
}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Second Level of Parallelization}
Version 3: Asynchronous, overlapping kernel calls

\begin{center}
\resizebox{!}{0.8\textheight}{%
\begin{tikzpicture}
	\tikzset{mywidth/.style={minimum width=2.5cm},};
	\node[block, mywidth] (l1) {\textbf{Worker 1}};
	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
										generate\\ nodes, \\
										start kernel
									\end{tabular}};
	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
	\node[frameless, mywidth, below=of l6] (l7) {};
	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
										receive\\ result, \\
										store nodes \\ 
										and edges
									\end{tabular}};
	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};

	\node[block, mywidth, right=of l1] (r1) {\textbf{Worker 2}};
	\node[frameless, mywidth, right=of l2, minimum height=1.5cm] (r2) {$\vdots$};
	\node[frameless, mywidth, right=of l3] (r3) {};
	\node[block, mywidth, right=of l6] (r6)	{\begin{tabular}{c}
										receive\\ result, \\
										store nodes \\
										and edges
									\end{tabular}};
	\node[block, mywidth, right=of l7] (r7) 	{\begin{tabular}{c}
										generate\\ nodes, \\
										start kernel
									\end{tabular}};
	\node[frameless, mywidth, right=of l12] (r12) {$\vdots$};

	\node[frameless, left=of l1] (g1) {};
	\node[block, below=of g1, minimum height=14cm] (graph) {\rotatebox{90}{\textbf{Graph}}};

	%gpu column
	\tikzset{gpuwidth/.style={minimum width=0.8cm},};

	\node[block, right=2cm of r1, minimum width=2.5cm] {\textbf{GPU}};
	\node[frameless, right=2cm of r2, gpuwidth] (topa) {$\vdots$};
	\node[frameless, right=of topa, gpuwidth] (topb) {$\vdots$};
	\node[block2, below=0.7cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=4.2cm] (stream1b) {\rotatebox{90}{kernel}};
	\node[frameless, below=0.3cm of stream1b, gpuwidth] (bota) {$\vdots$};

	\node[block2, below=2.2cm of topb, gpuwidth, minimum height=7.6cm] (stream2) {\rotatebox{90}{kernel}};
	\node[frameless, right=of bota, gpuwidth] (botb) {$\vdots$};

	%EDGES

	\path[-latex',thick]	(graph.east |- l3.west) edge (l3.west)
					(l11.west) edge (graph.east |- l11.west);
	\path[-latex',thick]	(graph.east |- r7.west) edge (r7.west)
					(r6.west) edge (graph.east |- r6.west);
	\path[dotted,thick]	(l2) edge (l3)
					(l3) edge (l11)
					(l11) edge (l12)
					(r2) edge (r6)
					(r6) edge (r7)
					(r7) edge (r12);

	\path[-latex',thick]	(r6.east -| stream1b.west) edge (r6.east);
	\path[-latex',thick]	(r7.east) edge (r7.east -| stream1b.west);

	\path[-,thick]		(l3.east) edge (l3.east -| stream1a.west);
	\path[-latex',thick]	(stream1a.east |- l3.east) edge (l3.east -| stream2.west);
	\path[latex'-,thick]		(l11.east) edge (l11.east -| stream1b.west);
	\path[-,thick]		(stream1b.east |- l11.east) edge (l11.east -| stream2.west);

\end{tikzpicture}
}
\end{center}

\end{frame}

\begin{frame}
\frametitle{Implementation and Testing}
	\begin{itemize}
	\item GPU imlementation with CUDA
	\item CPU Parallelization with MPI
	\item Tested with a 4 axis robot scenario created with CAD/Matlab
	\end{itemize}
	\begin{center}
	\includegraphics[height=4cm]{robotsolution.pdf}
	\end{center}
\end{frame}

\begin{frame}
\frametitle{Literatur}
\bibliographystyle{alpha}
\begin{thebibliography}{999}
	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, source: http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html
	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press, 2005
	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press, 2006
	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München, 2014
	\bibitem{chung} K. Chung, Wenping Wang, Quick Collision Detection of Polytopes in Virtual Environments, Proceedings of the ACM Symposium on Virtual Reality Software and Technology (VRST 96), Mark Green (ed.), 1996
	\bibitem{ericson} C. Ericson, Real-Time Collision Detection, Elsevier, 2005
\end{thebibliography}
\end{frame}

\end{document}